### 2021-02-21 architecture.tex
Describe ranking (symbol cost to precedence)

\subsection{Sort: From symbol costs to precedence}

>>>>

\subsection{Ranking: From symbol costs to precedence}

We extend the notion of cost from symbols to precedences
by summing up the symbol costs
weighted by their positions in the given precedence $\pi$:
$$
C(\pi) = \sum_{i=1}^n c(\pi(i)) \cdot i
$$

\begin{lemma}
Precedence cost $C$ is minimized by the precedence that ranks the symbols by their costs in non-increasing order:
% FB: The ranking terminology is from Fast Differentiable [...] Ranking.
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
\end{lemma}

\begin{proof}
We prove the lemma by contradiction.
Let $\pi$ minimize $C$ and let $\pi$ not rank the costs in non-increasing order.
Then there exist $k < l$ such that $c(\pi(k)) < c(\pi(l))$.
Let $\bar{\pi}$ be a precedence obtained from $\pi$ by swapping the elements $k$ and $l$.
%$\bar{\pi} = (\pi(1), \ldots, \pi(k-1), \pi(l), \pi(k+1), \ldots, \pi(l-1), \pi(k), \pi(l+1), \ldots, \pi(n))$.
\begin{align*}
C(\bar{\pi}) - C(\pi)
&= kc(\bar{\pi}(k)) + lc(\bar{\pi}(l)) - kc(\pi(k)) - lc(\pi(l)) \\
&= kc(\pi(l)) + lc(\pi(k)) - kc(\pi(k)) - lc(\pi(l)) \\
&= k(c(\pi(l)) - c(\pi(k))) - l(c(\pi(l)) - c(\pi(k))) \\
&= (k-l) (c(\pi(l)) - c(\pi(k))) \\
&< 0
\end{align*}
The final inequality is due to the assumption that $k-l < 0$ and
$c(\pi(l)) - c(\pi(k)) > 0$.
Thus, $C(\bar{\pi}) < C(\pi)$, which is contradicts the assumption that $\pi$ minimizes $C$. \qed
\end{proof}

Once the symbol costs are known,
producing a precedence that minimizes the precedence cost is cheap
because ranking is a fast operation.

Note that the precedence that minimizes $C$
orders the symbols with the lowest cost as the last,
prioritizing them for early inferences.

\todo[inline]{FB: Why is the equality predicate always the first in the precedence?}
